---
editor_options:
  markdown:
    wrap: 72
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Probability and Statistics

# Lab Assignment 1: Naive Bayes Classifier

### *Ivan Varanytsia, Yaryna Fialko*

## Introduction

During the past three weeks, you learned a couple of essential notions
and theorems, and one of the most important among them is the *Bayes
theorem*.

One of its applications is **Naive Bayes classifier**, which is a
probabilistic classifier whose aim is to determine which class some
observation probably belongs to by using the Bayes formula:
$$\mathsf{P}(\mathrm{class}\mid \mathrm{observation})=\frac{\mathsf{P}(\mathrm{observation}\mid\mathrm{class})\mathsf{P}(\mathrm{class})}{\mathsf{P}(\mathrm{observation})}$$

Under the strong independence assumption, one can calculate
$\mathsf{P}(\mathrm{observation} \mid \mathrm{class})$ as
$$\mathsf{P}(\mathrm{observation}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i), \qquad \mathsf{P}(\mathrm{observation} \mid \mathrm{class}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i \mid \mathrm{class}),$$
where $n$ is the total number of features describing a given
observation. Thus, $\mathsf{P}(\mathrm{class}|\mathrm{observation})$ now
can be calculated as

$$\mathsf{P}(\mathrm{class} \mid \mathrm{\mathrm{observation}}) = \mathsf{P}(\mathrm{class})\times \prod_{i=1}^{n}\frac{\mathsf{P}(\mathrm{feature}_i\mid \mathrm{class})}{\mathsf{P}(\mathrm{feature}_i)}\tag{1}$$

All the terms on the right-hand side can be estimated from the data as
respective relative frequencies;\
see [this
site](https://monkeylearn.com/blog/practical-explanation-naive-bayes-classifier/)
for more detailed explanations.

## Data description

### 11-th team = 1 (mod 5)

-   **1 - discrimination** This data set consists of tweets that have
    discriminatory (sexism or racism) messages or of tweets that are of
    neutral mood. The task is to determine whether a given tweet has
    discriminatory mood or does not.

# *train.csv* is used to build a bag-of-words, while *test.csv* - to see how well the Bayes Classifier works


```{r Including packages}
# here goes a list of recommended libraries,
# though you may install other ones if they are needed


library(tidytext)
library(readr)
library(dplyr)
library(devtools)
devtools::install_github("lchiffon/wordcloud2")
library(ggplot2)
library(RColorBrewer)
library(wordcloud2)
```

## Instructions

-   The first step is data pre-processing, which includes removing
    punctuation marks and stop words

-   represent each message as a bag-of-words

-   using the training set, calculate all the conditional probabilities
    in formula (1)

-   use those to predict classes for messages in the test set

-   evaluate effectiveness of the classifier by calculating the
    corresponding metrics

-   shortly summarize your work

-   do not forget to submit both the (compiled) Rmd source file and the .html
    output

### Data pre-processing

-   Read the *.csv* data files.
-   Сlear your data from punctuation or other unneeded symbols.
-   Clear you data from stop words. You don't want words as is, and, or
    etc. to affect your probabilities distributions, so it is a wise
    decision to get rid of them. Find list of stop words in the cms
    under the lab task.
-   Represent each test message as its bag-of-words. Here:
    <https://machinelearningmastery.com/gentle-introduction-bag-words-model/>
    you can find general introduction to the bag-of-words model and
    examples on to create it.
-   It is highly recommended to get familiar with R dataframes, it would
    make the work much easier to do.
-   Useful links:
    -   <https://steviep42.github.io/webscraping/book/bagofwords.html#tidytext> -
        example of using *tidytext* to count frequencies of the words.
    -   Basics of Text Mining in R:
        <http://rstudio-pubs-static.s3.amazonaws.com/256588_57b585da6c054349825cba46685d8464.html>
        . Note that it also includes an example on how to create a bag
        of words from your text document.




## Linking all needed files and parsing stop words:

```{r Adding stop words}
test_path <- "data/1-discrimination/test.csv"
train_path <- "data/1-discrimination/train.csv"

stop_words <- read_file("stop_words.txt")
splitted_stop_words <- strsplit(stop_words, split = '\n')
splitted_stop_words <- splitted_stop_words[[1]]
```
## Reading files:
```{r Reading the data}
train <- read.csv(file = train_path, stringsAsFactors = FALSE)
test <- read.csv(file = test_path, stringsAsFactors = FALSE)
```

```{r Cleaning the data}
# note the power functional features of R bring us!
tidy_text <- unnest_tokens(train, output = 'word', input = 'tweet') %>%
  filter(!word %in% splitted_stop_words)

dataframe <- tidy_text %>% count(word, sort = TRUE)

df_labels <- tidy_text %>% count(label, word, sort = TRUE)
df_discrim <- df_labels %>% filter(label == "discrim")
df_neutral <- df_labels %>% filter(label == "neutral")
```

### Data visualization

Each time you work with some data, you need to understand it before you
start processing it. R has very powerful tools to make nice plots and
visualization. Show what are the most common words for negative and
positive examples as a histogram, word cloud etc. Be creative!


### TODO - Visualize the most frequent words


```{r Helping values}
discrim_test = test %>% filter(label == "discrim") %>% count(label, tweet, result = 1)
neutral_test = test %>% filter(label == "neutral") %>% count(label, tweet, result = 1)
is_discrim = nrow(train %>% filter(label == "discrim")) / nrow(train)
is_neutral = 1 - is_discrim
```


```{r Class Definition}
naiveBayes <- setRefClass("naiveBayes",

                          # d_test - for discrimination tweets
                          # n_test - for neutral tweets
                          fields = list(d_test = "data.frame", n_test = "data.frame"),

                          methods = list(
                            fit = function()
                            {
                              for (row in 1:nrow(neutral_test)) {
                                value = predict(neutral_test[row, ])
                                  if (value == 0){
                                  neutral_test[row, "result"] = 0
                                  }
                              }
                              for (row in 1:nrow(discrim_test)) {
                                value = predict(discrim_test[row, ])
                                 if(value == 1){             
                                  discrim_test[row, "result"] = 0
                               }
                              }
                              d_test <<- discrim_test
                              n_test <<- neutral_test
                            },


                            # returns prediction for a tweet 0 - discriminative, 1 - neutral
                            predict = function(tweet)
                            {
                              words <- unnest_tokens(tweet, output = 'word', input = 'tweet')
                              dfs <- list(df_discrim, df_neutral)
                              result <- list(0, 0)
                              
                              current_df <- dfs[[1]]
                              cond_prob <- 1
                              
                              for (row in 1:nrow(words)) {
                                  cond_prob <- cond_prob * probab(words[row, "word"], current_df)
                              }
                              result[[1]] = cond_prob * is_discrim
                              current_df <- dfs[[2]]
                              cond_prob <- 1
                              
                              for (row in 1:nrow(words)) {
                                  cond_prob <- cond_prob * probab(words[row, "word"], current_df)
                              }
                              result[[2]] = cond_prob * is_neutral
                              
                              if (result[[1]] > result[[2]]) {
                                return(0)
                              }
                              return(1)
                            },
                            
                            probab = function(word, current_df) {
                              p <- which(current_df["word"] == word)
                              denominator <- nrow(current_df) + nrow(df)
                              if (length(p) == 0){
                                  return(1 / denominator)
                              }
                              else {
                                  return ((1 + current_df["n"][p, ]) / denominator)
                              }
                            },

                            score = function()
                            {
                              neutral_f <- n_test[n_test$result == FALSE,]
                              neutral_f_amount <- nrow(neutral_f)
                              neutral_t_amount = nrow(n_test) - nrow(neutral_f)

                              dicrim_t <- d_test[d_test$result == TRUE,]
                              discrim_t_amount <- nrow(dicrim_t)
                              discrim_f_amount = nrow(d_test) - nrow(dicrim_t)


                              precision <- neutral_t_amount / (neutral_t_amount +
                                discrim_f_amount)
                              recall <- neutral_t_amount / (neutral_f_amount+neutral_t_amount)
                              F1 <- 2 * (precision * recall / (precision + recall))


                              variants <- c(discrim_t_amount, neutral_t_amount, discrim_f_amount, neutral_f_amount)
                              par(mfrow = c(1, 2))
                              myPalette <- brewer.pal(5, "Set2")


                              pie(variants, labels=c("correct discrim ", "correct neutral",
                                                    "incorrect discrim", "incorrect neutral"), border="black", col=myPalette )


                              accuracy <- (discrim_t_amount + neutral_t_amount) / (discrim_f_amount + discrim_t_amount + neutral_f_amount + neutral_t_amount)
                              mtrcs <- c(precision, recall, F1, accuracy)
                              barplot(
                                mtrcs,
                                main = "Predictions acuraccy",
                                xlab = "Metrics",
                                ylab = "Values",
                                ylim = c(0, 1),
                                names.arg = c("Precision", "Sensitivity", "F1", "Accuracy"),
                                col = c("blue", "blue", "yellow", "yellow"),
                                horiz = FALSE)
                            },
                            
                            frequencyofwords = function()
                              {
                                par(mfrow = c(1, 3))
                              
                                
                              wordcloud(words = df_discrim$word, freq = df_discrim$n, min.freq = 3, max.words=100, random.order=FALSE,rot.per=0.35, colors=brewer.pal(8, "Dark2"), use.r.layout=FALSE)
                                #wordcloud2(data=df_discrim$word, size=1.6, color='random-dark')
                                plot.new()
                                wordcloud(words = df_neutral$word, freq = df_neutral$n, min.freq = 3, max.words=100, random.order=FALSE,rot.per=0.35, colors=brewer.pal(8, "Dark2"), use.r.layout=FALSE)
                                #wordcloud2(data=df_neutral$word, size=1.6, color='random-dark')

                      
                            }
                          ))

run = naiveBayes()
run$fit()
run$score()
run$frequencyofwords()

```

## Measure effectiveness of your classifier
-   Note that accuracy is not always a good metric for your classifier.
-   Look at precision and recall curves, F1 score metric.
-   Visualize them.
-   Show failure cases.
## Conclusions

Summarize your work by explaining in a few sentences the points listed
below.

-   Describe the method implemented in general. Show what are
    mathematical foundations you are basing your solution on.
-   List pros and cons of the method. This should include the
    limitations of your method, all the assumption you make about the
    nature of your data etc.